---
title: "Logistic Regression on Byssinosis Data"
author: "Christina Chang (ID:913970829)"
date: "12/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE)
```

```{r}
# Import libraries.


# Read the data.
df = read.csv("byssinosis.csv")
```


This project will investigate the relationship between the Byssinosis disease and smoking status, sex, race, length of employment, smoking, and dustiness of the workplace.


#### Model fitting
```{r}
model = glm(formula = cbind(Byssinosis,Non.Byssinosis) ~ ., 
            family = binomial(), data = df)
summary(model)
```
##### Interpreting logistic regression results
From the results of the model, sex and race are not statistically significant. As for the statistically significant variables, Workspace has the smallest p-value, therefore, there is a strong association between the type of work place and the disease. Since the coefficient for Workspace is negative, this means that an environment that is less dusty decreases the odds of having the disease, given that all other variables are held constant.

```{r}
# Consider interaction terms.
model = glm(formula = cbind(Byssinosis,Non.Byssinosis) ~ (.)^2, 
            family = binomial(), data = df)
summary(model)
```


#### Perform stepwise variable selection
```{r}
fullmodel = glm(formula = cbind(Byssinosis,Non.Byssinosis) ~ ., 
            family = binomial(), data = df)
nullmodel = glm(formula = cbind(Byssinosis,Non.Byssinosis) ~ 1, 
            family = binomial(), data = df)
```



#### Appendix
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE,include =TRUE}
```